Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-06-16 omrp_lb_rev.tex
and & or

Overlapping batch means modifies this idea by taking batches given by $\yb_j(m) = \frac{1}{m} \sum_{i=1}^{m}$ $y^{(j-1)+i}$, for $j = 1, \dots, n-m+1$.
As an example, suppose each batch contains $m$ observations and that there are a total of $n=2m$ observations.
In the nonoverlapping case, there are only two batches: the first batch contains observations $y^1, y^2, \dots, y^m$ and the second batch contains observations $y^{m+1}, y^{m+2}, \dots, y^{2m}$.   
In the overlapping batches of \citet{Meketon1984}, the first batch is as before, the second batch now consists of $y^2, y^3, \dots, y^{m+1}$, the third batch contains observations $y^3, y^4, \dots, y^{m+2}$ and so on.  
This is what we call the \emph{maximally-overlapping} batch means method, because $m-1$ observations are common to---and only one observation changes between---adjacent batches.

@@

As before, let $m$ denote the batch size, $n$ the total sample size and $k = \lfloor \tfrac{n}{m} \rfloor$ be the number of nonoverlapping batches.  
In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch estimator is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
where $\nb$ is the number of batches used given $n$, $m$ and $\gamma$.  
The sample variance estimator given in (\ref{eq:overlap_variance_def}) is changed to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}
Note that with $\gamma=1$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:overlap_variance_def}) and with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).

@@

In order to apply overlapping batches to MRP, we need to keep track of solutions to sampling problems (SP$_m$) for each batch of size $m$.  
Toward this end, let $B(i)$ denote the set of batches $j \in \left\{1, 2, \dots, \nb \right\}$ observation $\xiti$ is used in, $i = 1, 2, \dots, n$.  
See Figure \ref{fig:overlap_nonint} for an example with $n=18$, $m = 6$ and $\gamma = 2$.  
Here, the first batch uses observations $\xit^1, \xit^2, \xit^3, \xit^4, \xit^5, \xit^6$, the second batch uses $\xit^3, \xit^4, \xit^5, \xit^6, \xit^7, \xit^8$, and so on.  
So, $B(1) = \{1\}$, $B(3) = \{1,2\}$, and $B(7)=\{2,3,4\}$.  
As before, we use $\xs_j$ to denote an optimal solution to sampling problem (SP$_m$) formed using the $j$th batch; $\xs_j \in \argmin_{\x \in X} \frac{1}{m} \sum_{i=1}^m f(\x,\xit^{\gamma(j-1) + i})$, $j = 1, 2, \dots, \nb$.

\begin{figure}[htb!]
	\centering
	\begin{tabular}{*{18}{c}}
		$\xit^1$ & $\xit^2$ & $\xit^3$ & $\xit^4$ & $\xit^5$ & $\xit^6$ & $\xit^7$ & $\xit^8$ & $\xit^9$ & $\xit^{10}$ & $\xit^{11}$ & $\xit^{12}$ & $\xit^{13}$ & $\xit^{14}$ & $\xit^{15}$ & $\xit^{16}$ & $\xit^{17}$  & $\xit^{18}$ \\
% 		First Set of Batches
		\multicolumn{6}{l}{[---------------1-----------------]} &
		\multicolumn{6}{l}{[-----------------4------------------]} &
		\multicolumn{6}{l}{[------------------7--------------------]} \\
% 		Third set of Batches
		& & \multicolumn{6}{l}{[---------------2-----------------]} &
		\multicolumn{6}{l}{[------------------5--------------------]} \\
% 		Fifth set of Batches
		& & & & \multicolumn{6}{l}{[---------------3-----------------]} &
		\multicolumn{6}{l}{[------------------6--------------------]} \\
	\end{tabular}
	\caption{Visual representation of overlapping batches with $n = 18$, $m = 6$ and $\gamma = 2$.  
        The brackets show which observations are used in each batch, and the numbers inside each bracket show the batch number $j$.}
	\label{fig:overlap_nonint}
\end{figure}

The results on overlapping batch means occur in the limit as $n, m, k=n/m \rightarrow \infty$ \citep{damerdji1994strong,damerdji1995mean,Meketon1984,Song1992,Welch1987}.  
Let $n_l, m_l$ and $k_l$ be sequences of numbers satisfying these requirements, then the limits will be taken as $l \rightarrow \infty$.  
We assume a batching structure where $m_l = \lfloor n_l ^r \rfloor$ for some $0<r<1$ as is typical in simulation, and $n_l$ tends to infinity as $l \rightarrow \infty$.  
In addition, we may desire that the batch nonoverlap parameter change with $l$, to ensure that $\gamma_l / m_l = \gammab_l$ converges to the constant $\gammab$.  
Now we are ready to define the estimators for OMRP.
\begin{align}
	\gb_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs_j,\xit^{\gamma_l(j-1)+1}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:gbar} \\
	\gbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \frac{1}{|B(i)|} \sum_{j \in B(i)} \left[ f(\xh,\xiti) - f(\xs_j,\xiti) \right], \label{eq:gbb} \\
	VG_l & = \dfrac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\gb_{lj} - \gbb_l)^2. \label{eq:vg}
\end{align}

@@

\begin{figure}[htb!]
	\centering
	\subfloat[][NVD]{
	   \label{fig:nv_varvar}
	   \includegraphics[width=.5\textwidth]{images/nv_varvar_discrete}}
	\subfloat[][CEP1]{
	   \label{fig:cep1_varvar}
	   \includegraphics[width=.5\textwidth]{images/cep1_varvar}}
	\caption{
		$\var{n VG}$ relative to the nonoverlapping counterpart across various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches) for test problem
		(a) NVD, and
		(b) CEP1
	}
\label{fig:varvar1}
\end{figure}

\begin{figure}[htb!]
	\centering
	\subfloat[][10D]{
	   \label{fig:10d_varvar}
	   \includegraphics[width=.5\textwidth]{images/10D_varvar}}
	\subfloat[][DB1]{
	   \label{fig:db1_varvar}
	   \includegraphics[width=.5\textwidth]{images/db1_varvar}}
		\caption{
		$\var{n VG}$ relative to the nonoverlapping counterpart across various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches) for test problem
		(a) 10D, and
		(b) DB1
		}
\label{fig:varvar2}
\end{figure}

\begin{figure}[htb!]
	\centering
	\subfloat[][Expected Variance]{
	   \label{fig:nv_avgvar}
	   \includegraphics[width=.5\textwidth]{images/nv_avgvar_discrete}}
	\subfloat[][Coverage Probability]{
	   \label{fig:nv_cover}
	   \includegraphics[width=.5\textwidth]{images/nv_cover_discrete}}
	\caption{ 
		Estimates of
		(a) $\e{nVG}$, and 
		(c) coverage probability of the CIs for various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches)
		 for NVD
	}
\label{fig:nv}
\end{figure}

>>>>

Overlapping batch means modifies this idea by taking batches given by $\yb_j(m) = \frac{1}{m} \sum_{i=1}^{m}$ $y^{(j-1)+i}$, for $j = 1, \dots, n-m+1$.
As an example, suppose each batch contains $m$ observations and that there are a total of $n=2m$ observations.
In the nonoverlapping case, there are only two batches: the first batch contains observations $y^1, y^2, \dots, y^m$ and the second batch contains observations $y^{m+1}, y^{m+2}, \dots, y^{2m}$.   
In the overlapping batches of \citet{Meketon1984}, the first batch is as before, the second batch now consists of $y^2, y^3, \dots, y^{m+1}$, the third batch contains observations $y^3, y^4, \dots, y^{m+2}$, and so on.  
This is what we call the \emph{maximally-overlapping} batch means method, because $m-1$ observations are common to---and only one observation changes between---adjacent batches.

@@

As before, let $m$ denote the batch size, $n$ the total sample size and $k = \lfloor \tfrac{n}{m} \rfloor$ be the number of nonoverlapping batches.  
In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch estimator is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
where $\nb$ is the number of batches used given $n$, $m$, and $\gamma$.  
The sample variance estimator given in (\ref{eq:overlap_variance_def}) is changed to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}
Note that with $\gamma=1$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:overlap_variance_def}) and with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).

@@

In order to apply overlapping batches to MRP, we need to keep track of solutions to sampling problems (SP$_m$) for each batch of size $m$.  
Toward this end, let $B(i)$ denote the set of batches $j \in \left\{1, 2, \dots, \nb \right\}$ observation $\xiti$ is used in, $i = 1, 2, \dots, n$.  
See Figure \ref{fig:overlap_nonint} for an example with $n=18$, $m = 6$, and $\gamma = 2$.  
Here, the first batch uses observations $\xit^1, \xit^2, \xit^3, \xit^4, \xit^5, \xit^6$, the second batch uses $\xit^3, \xit^4, \xit^5, \xit^6, \xit^7, \xit^8$, and so on.  
So, $B(1) = \{1\}$, $B(3) = \{1,2\}$, and $B(7)=\{2,3,4\}$.  
As before, we use $\xs_j$ to denote an optimal solution to sampling problem (SP$_m$) formed using the $j$th batch; $\xs_j \in \argmin_{\x \in X} \frac{1}{m} \sum_{i=1}^m f(\x,\xit^{\gamma(j-1) + i})$, $j = 1, 2, \dots, \nb$.

\begin{figure}[htb!]
	\centering
	\begin{tabular}{*{18}{c}}
		$\xit^1$ & $\xit^2$ & $\xit^3$ & $\xit^4$ & $\xit^5$ & $\xit^6$ & $\xit^7$ & $\xit^8$ & $\xit^9$ & $\xit^{10}$ & $\xit^{11}$ & $\xit^{12}$ & $\xit^{13}$ & $\xit^{14}$ & $\xit^{15}$ & $\xit^{16}$ & $\xit^{17}$  & $\xit^{18}$ \\
% 		First Set of Batches
		\multicolumn{6}{l}{[---------------1-----------------]} &
		\multicolumn{6}{l}{[-----------------4------------------]} &
		\multicolumn{6}{l}{[------------------7--------------------]} \\
% 		Third set of Batches
		& & \multicolumn{6}{l}{[---------------2-----------------]} &
		\multicolumn{6}{l}{[------------------5--------------------]} \\
% 		Fifth set of Batches
		& & & & \multicolumn{6}{l}{[---------------3-----------------]} &
		\multicolumn{6}{l}{[------------------6--------------------]} \\
	\end{tabular}
	\caption{Visual representation of overlapping batches with $n = 18$, $m = 6$, and $\gamma = 2$.  
        The brackets show which observations are used in each batch, and the numbers inside each bracket show the batch number $j$.}
	\label{fig:overlap_nonint}
\end{figure}

The results on overlapping batch means occur in the limit as $n, m, k=n/m \rightarrow \infty$ \citep{damerdji1994strong,damerdji1995mean,Meketon1984,Song1992,Welch1987}.  
Let $n_l, m_l$, and $k_l$ be sequences of numbers satisfying these requirements, then the limits will be taken as $l \rightarrow \infty$.  
We assume a batching structure where $m_l = \lfloor n_l ^r \rfloor$ for some $0<r<1$, as is typical in simulation, and $n_l$ tends to infinity as $l \rightarrow \infty$.  
In addition, we may desire that the batch nonoverlap parameter change with $l$, to ensure that $\gamma_l / m_l = \gammab_l$ converges to the constant $\gammab$.  
Now we are ready to define the estimators for OMRP.
\begin{align}
	\gb_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs_j,\xit^{\gamma_l(j-1)+1}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:gbar} \\
	\gbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \frac{1}{|B(i)|} \sum_{j \in B(i)} \left[ f(\xh,\xiti) - f(\xs_j,\xiti) \right], \label{eq:gbb} \\
	VG_l & = \dfrac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\gb_{lj} - \gbb_l)^2. \label{eq:vg}
\end{align}

@@

\begin{figure}[htb!]
	\centering
	\subfloat[][NVD]{
	   \label{fig:nv_varvar}
	   \includegraphics[width=.5\textwidth]{images/nv_varvar_discrete}}
	\subfloat[][CEP1]{
	   \label{fig:cep1_varvar}
	   \includegraphics[width=.5\textwidth]{images/cep1_varvar}}
	\caption{
		$\var{n VG}$ relative to the nonoverlapping counterpart across various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches) for test problems
		(a) NVD and
		(b) CEP1
	}
\label{fig:varvar1}
\end{figure}

\begin{figure}[htb!]
	\centering
	\subfloat[][10D]{
	   \label{fig:10d_varvar}
	   \includegraphics[width=.5\textwidth]{images/10D_varvar}}
	\subfloat[][DB1]{
	   \label{fig:db1_varvar}
	   \includegraphics[width=.5\textwidth]{images/db1_varvar}}
		\caption{
		$\var{n VG}$ relative to the nonoverlapping counterpart across various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches) for test problems
		(a) 10D and
		(b) DB1
		}
\label{fig:varvar2}
\end{figure}

\begin{figure}[htb!]
	\centering
	\subfloat[][Expected Variance]{
	   \label{fig:nv_avgvar}
	   \includegraphics[width=.5\textwidth]{images/nv_avgvar_discrete}}
	\subfloat[][Coverage Probability]{
	   \label{fig:nv_cover}
	   \includegraphics[width=.5\textwidth]{images/nv_cover_discrete}}
	\caption{ 
		Estimates of
		(a) $\e{nVG}$ and 
		(b) coverage probability of the CIs for various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches)
		 for NVD
	}
\label{fig:nv}
\end{figure}
